# 4.3.2020

# This script will show 
1. the shift towards positive estimates for the association between gene expression and mtDNA-CN (both blood)

2. A table showing the most significant genes

3. A table showing significance for transcription factors

#### load data
```{r}
load('projects/GTeX/syang/look.version8/Whole Blood/with.gene.rds')

load('/Volumes/JHPCE/dcs01/resources/ensembl/only.care.rds') # made in /dcs01/arking/arkinglab/resources/ensembl/try.read.R
w.annot <- merge(only.care, with.gene, by = 'gene_id') # 23758 yay
w.annot$symbol <- w.annot$symbol.y
```

#### subset no pseudogenes
```{r}
gene.types <- unique(w.annot$gene_type)
pseudo.types <- gene.types[grep('pseudo', gene.types)]
'%!in%' <- function(x,y)!('%in%'(x,y))
with.gene <- subset(w.annot, gene_type %!in% pseudo.types)
```

# This will show a shift towards positively associated genes

# 1. plot distribution
```{r}
load('projects/GTeX/syang/look.version8/Whole Blood/with.gene.rds')

library(ggplot2)
ggplot(with.gene, aes(t_value)) + geom_rug() + geom_density() 
sig <- subset(with.gene, pval < 0.05)

ggplot(sig, aes(t_value)) + geom_rug() + geom_density(color = 'black', fill = '#999999') + xlab('T-value') + theme_classic()

sig$dir <- ifelse(sig$beta < 0, 'Neg', 'Pos')
table(sig$dir)
```

# 2. make table
```{r}
library(htmlTable)
with.gene <- with.gene[order(with.gene$pval),]
top.ten <- subset(with.gene, pval < 0.05)
top.ten <- dplyr::select(top.ten, symbol, gene_type, beta, pval)

top.ten$beta <- formatC(top.ten$beta, format = "e", digits = 2)
top.ten$pval <-ifelse(top.ten$pval > 0.001, format(round(top.ten$pval, 3), nsmall = 3)
, formatC(top.ten$pval, format = "e", digits = 2))

top.ten$gene_type <- gsub('protein_coding', 'protein coding', top.ten$gene_type)

only.matter <- dplyr::select(with.gene, symbol, gene_type, beta, pval)

library(htmlTable)
# htmlTable(only.matter,
#           header =  c('Gene', '&emsp;&emsp;Gene Class&emsp;&emsp;', '&emsp;&emsp;&emsp;&emsp;Effect size&emsp;&emsp;&emsp;&emsp;', '&emsp;&emsp;P-value&emsp;&emsp;'),
#           rnames = rep('', nrow(top.ten)))
```

# 2.5 get annotations for genes
```{r}
# read in all sets for go.
library(qusage)
# go.sets <- read.gmt('/Volumes/JHPCE 1/dcs01//active/projects/GTeX/syang/look.version8/GO_enrich/MsigDB/c5.all.v7.0.symbols.gmt')

# GO is far too slow. use KEGG.

kegg.sets <- read.gmt('/Volumes/JHPCE 1/dcs01/active/projects/GTeX/syang/look.version8/GO_enrich/MsigDB/c2.cp.kegg.v7.0.symbols.gmt')

# certain things to highlight
# here are ones that are across tissue.....
# 
# carry.annot <- as.data.frame(matrix(nrow = 1, ncol = 2))
# colnames(carry.annot) <- c('Gene', 'Annotation')
# 
# for(i in 1:nrow(top.ten)){
#   # exact searches do not work!
#   # to.search <- paste0('^', top.ten$symbol[i], '$')
#   to.search <- paste0(top.ten$symbol[i])
#   indices <- grep(to.search, kegg.sets)
#   kegg.terms <- names(kegg.sets[indices])
#   to.add <- data.frame(Gene = rep(to.search, length(kegg.terms)), Annotation = kegg.terms)
#   carry.annot <- rbind(carry.annot, to.add)
#   if(i %% 100 == 0){
#     print(paste('On gene', i))
#   }
# }
# save(carry.annot, file = '/Volumes/JHPCE 1/dcs01/active/projects/GTeX/syang/look.version8/Whole Blood/carry.annot.rds')
load('projects/GTeX/syang/look.version8/Whole Blood/carry.annot.rds')
```

# 2.75 check
```{r}

freq <- as.data.frame(table(carry.annot$Annotation))
freq <- freq[order(freq$Freq, decreasing = T),]

head(freq, 10)

pathway1 <- 'KEGG_ENDOCYTOSIS'
pathway2 <- 'KEGG_MAPK_SIGNALING_PATHWAY'
check.overlap <- function(carry.annot, pathway1, pathway2){
  set1 <- subset(carry.annot, Annotation == pathway1)
  set2 <- subset(carry.annot, Annotation == pathway2)
  intersect(set1$Gene, set2$Gene) %>% length -> both
  print(paste(nrow(set1), nrow(set2), both))
}

check.overlap(carry.annot, 'KEGG_ENDOCYTOSIS', 'KEGG_MAPK_SIGNALING_PATHWAY')
check.overlap(carry.annot, 'KEGG_ENDOCYTOSIS', 'KEGG_UBIQUITIN_MEDIATED_PROTEOLYSIS')
check.overlap(carry.annot, 'KEGG_MAPK_SIGNALING_PATHWAY', 'KEGG_UBIQUITIN_MEDIATED_PROTEOLYSIS')

mapk <- subset(carry.annot, Annotation == 'KEGG_MAPK_SIGNALING_PATHWAY')
ubq <- subset(carry.annot, Annotation == 'KEGG_UBIQUITIN_MEDIATED_PROTEOLYSIS')
endo <- subset(carry.annot, Annotation == 'KEGG_ENDOCYTOSIS')


mapk2 <- only.matter[which(only.matter$symbol %in% mapk$Gene),]
ubq2 <- only.matter[which(only.matter$symbol %in% ubq$Gene),]
endo2 <- only.matter[which(only.matter$symbol %in% endo$Gene),]

# take only bonferroni significant:
mapk3 <- subset(mapk2, pval < 2.423772e-06) # 13
ubq3 <- subset(ubq2, pval < 2.423772e-06) # 13
endo3 <- subset(endo2, pval < 2.423772e-06) # 11

# that's still too many!  Just take the top 5. 
to.show <- rbind(head(mapk2, 5), head(endo2, 5), head(ubq2, 5))
```

### uhh...just use the top 3 KEGG terms.
```{r}
load('/Volumes/JHPCE 1/dcs01/active/projects/GTeX/syang/look.version8/Whole Blood/all.kegg.rds')

head(all.kegg.sets$Gene.Set.Name, 5)
cyt <- subset(carry.annot, Annotation == 'KEGG_CYTOKINE_CYTOKINE_RECEPTOR_INTERACTION')
ubq <- subset(carry.annot, Annotation == 'KEGG_UBIQUITIN_MEDIATED_PROTEOLYSIS')
splice <- subset(carry.annot, Annotation == 'KEGG_SPLICEOSOME')

cyt2 <- only.matter[which(only.matter$symbol %in% cyt$Gene),]
ubq2 <- only.matter[which(only.matter$symbol %in% ubq$Gene),]
splice2 <- only.matter[which(only.matter$symbol %in% splice$Gene),]

to.show <- rbind(head(cyt2, 5), head(ubq2, 5), head(splice2, 5))
```

### table
```{r}
to.show$beta <- formatC(to.show$beta, format = "e", digits = 2)
to.show$pval <-ifelse(to.show$pval > 0.001, format(round(to.show$pval, 3), nsmall = 3)
, formatC(to.show$pval, format = "e", digits = 2))
to.show <- dplyr::select(to.show, -gene_type)

htmlTable(to.show,
          header =  c('Gene', '&emsp;&emsp;&emsp;&emsp;Effect size&emsp;&emsp;&emsp;&emsp;', '&emsp;&emsp;P-value&emsp;&emsp;'),
          rnames = rep('', nrow(to.show)),
          rgroup = c('Cytokine-cytokine receptor interaction',
                     "Ubiquitin-mediated proteolysis", "Spliceosome"),
          n.rgroup = c(5,5,5))
```




# 3. TFs
sidetracked to tf land
the top 5 TFs: how do they look?
unclear is gabp b is gabpb1 or gabpb2, so have both.

```{r}
tfs <- rbind(only.matter[grep('ELK1', only.matter$symbol),], only.matter[grep('^NRF1$', only.matter$symbol),], only.matter[grep('^YY1$', only.matter$symbol),], only.matter[grep('^GABPB1$', only.matter$symbol),], only.matter[grep('^GABPB2$', only.matter$symbol),], only.matter[grep('^E4F1$', only.matter$symbol),])

tfs$beta <- formatC(tfs$beta, format = "e", digits = 2)
tfs$pval <-ifelse(tfs$pval > 0.001, format(round(tfs$pval, 3), nsmall = 3)
, formatC(tfs$pval, format = "e", digits = 2))
tfs <- dplyr::select(tfs, -gene_type)

library(htmlTable)
htmlTable(tfs,
          header =  c('Gene', '&emsp;&emsp;&emsp;&emsp;Effect size&emsp;&emsp;&emsp;&emsp;', '&emsp;&emsp;P-value&emsp;&emsp;'),
          rnames = rep('', nrow(tfs)))
```