# updated 4.20.2020
# 3.9.2020

# This script will look at GO pathways significant in multiple tissues! Adapted from projects/GTeX/syang/look.version8/GO_enrich/Test.GO.all.tissues/tabulate.cross.tissue.GO.nopseudo.R
```{r setup}
library(data.table)
library(knitr)
library(ggplot2)
library(magrittr)
wdir <- paste0('projects/GTeX/syang/look.version8/')
setwd(wdir)
opts_knit$set(root.dir = wdir)
```

# load results for tissues
```{r}
# load libraries
# install stuff
# library(devtools)
# source('/Volumes/JHPCE/dcs01/users/syang/libLoad.R')
# 
# # load full list of tissues
# load('projects/GTeX/syang/look.version8/R_objects/all_tissuesv8.rds')
# 
# load('projects/GTeX/syang/look.version8/Whole Blood/sig.pathways.nopseudo.permutecut.rds')
# big.sig.pathways <- all.results
# big.sig.pathways$Tissue <- 'Whole Blood'
# 
# all_tissues <- all_tissues[1:54]
# 
# for (i in 1:length(all_tissues)){
# 	char <- all_tissues[i]
# 	wd <- paste0('projects/GTeX/syang/look.version8/', char)
# 	if(file.exists(paste0(wd, '/sig.pathways.nopseudo.permutecut.rds'))){
# 		load(paste0(wd, '/sig.pathways.nopseudo.permutecut.rds'))
# 	  if(nrow(all.results) == 0){}else{
# 	  all.results$Tissue <- char}
# 		big.sig.pathways <- rbind.fill(big.sig.pathways, all.results)
# 	}
# }
# 
# save(big.sig.pathways, file = 'projects/GTeX/syang/look.version8/R_code/Cross.tissue.look/big.sig.pathways.nopseudo.permutecut.rds')

load('projects/GTeX/syang/look.version8/R_code/Cross.tissue.look/big.sig.pathways.permutecut_1000.rds')
```

# limit to only tissues with significant GIF
```{r}
# limit this so only looking at tissues with significant GIF
load('projects/GTeX/syang/look.version8/R_objects/lambda.frame.rds')
load('projects/GTeX/syang/look.version8/R_objects/all.lambdas.rds')
lambda.cutoff <- all.lambdas$Lambda[nrow(all.lambdas) * 0.05] 

sig.tissues <- subset(lambda.frame, Lambda > lambda.cutoff)

only.sig.tissues <- subset(big.sig.pathways, Tissue %in% sig.tissues$Tissue)
save(only.sig.tissues, file = 'projects/GTeX/syang/look.version8/R_code/Cross.tissue.look/only.sig.tissues_1000.rds')
```

# pathways significant in multiple tissues
# take out negative betas.
```{r}
only.sig.tissues$Beta <- as.numeric(only.sig.tissues$Beta)

only.sig.tissues = subset(only.sig.tissues, Beta > 0)
lookgenes2 <- as.data.frame(table(only.sig.tissues$Gene.Set.Name))
lookgenes2 <- lookgenes2[order(lookgenes2$Freq, decreasing = T),]

looktiss <- as.data.frame(table(only.sig.tissues$Tissue))
looktiss <- looktiss[order(looktiss$Freq, decreasing = T),]
View(lookgenes2)

high <- subset(lookgenes2, Freq > 13)

ggplot(lookgenes2, aes(Freq)) + geom_density() + geom_rug() + geom_vline(xintercept = high$Freq, lty = 2, colour = "gray50") + 
  geom_text(data = high, aes(x=high$Freq, y = .1), label = high$Var1, vjust = -0.5, hjust = 0, size = 4, angle = 90, colour = "gray50") + xlab('Number of tissues significant in')

set.info <- dplyr::select(only.sig.tissues, Gene.Set.Name, Set)
```

## distribution of lookgenes2
```{r}
# dan says use a barplot
# ggplot(lookgenes2, aes(Freq)) + geom_density() + geom_rug()

# Basic barplot
top20 = head(lookgenes2, 10)

top20$Var1 = factor(top20$Var1, levels = top20$Var1[order(top20$Freq, decreasing = T)])

p<-ggplot(top20, aes(x=Var1, y=Freq)) + geom_bar(stat="identity")

p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## show lookgenes2 prettily
```{r}
w.info <- merge(lookgenes2, set.info, by.x = 'Var1', by.y = 'Gene.Set.Name')

u <- unique(w.info)
u <- u[order(u$Freq, decreasing = T),]
no.tfs <- subset(u, Set != "TFT")
tfs <- subset(u, Set == "TFT")
mito <- u[grep('MITOC', u$Var1),]
kegg <- subset(u, Set == 'KEGG')

show <- rbind(head(tfs, 5), head(no.tfs, 5), head(kegg,5), head(mito,5))
show <- dplyr::select(show, Var1, Freq)

library(htmlTable)
htmlTable(show,
          header =  c('Pathway', '&emsp;&emsp;&emsp;&emsp;Number of significant tissues&emsp;&emsp;&emsp;&emsp;'),
          rnames = rep('', nrow(lookgenes2)),
          rgroup = c('Transcription factors', "GO terms", 'KEGG terms', "Mitochondrial terms"),
          n.rgroup = c(5,5,5,5))
# 
# show$Var1 <- tolower(show$Var1)
# show$Var1 <- gsub('go_', '', show$Var1)
# show$Var1 <- gsub('kegg_', '', show$Var1)
```
