# 5.15.2020
# This script will look at drivers for KEGG pathways
# Also ELK1/other stuff
# Adapted from Meta.Forestplots.Rmd

# 3.6.2020

# This script will show random beta meta-analysis results. Adapted from Cross.tissue.look/forestplots.R
```{r setup}
library(data.table)
library(knitr)
library(meta)
wdir <- paste0('/Volumes/JHPCE/dcs01/active/projects/GTeX/syang/look.version8/')
setwd(wdir)
opts_knit$set(root.dir = wdir)
```

# get Ns for each!
```{r}
load('/Volumes/JHPCE/dcs01/active/projects/GTeX/syang/look.version8/R_objects/big.with.gene.w.samps.rds')

load('R_objects/all.lambdas.rds')
load('R_objects/lambda.frame.rds') 

sig <- subset(lambda.frame, Lambda > all.lambdas$Lambda[nrow(all.lambdas)*0.05])

only.sig <- subset(w.samps, tissue %in% sig$Tissue)
```

# What genes are driving HD?
```{r}
library(qusage)
kegg.sets <- read.gmt('/Volumes/JHPCE/dcs01/active/projects/GTeX/syang/look.version8/GO_enrich/MsigDB/c2.cp.kegg.v7.0.symbols.gmt')
tft.sets <- read.gmt('/Volumes/JHPCE/dcs01/active/projects/GTeX/syang/look.version8/GO_enrich/MsigDB/c3.tft.v7.0.symbols.gmt')
go.sets <- read.gmt('/Volumes/JHPCE/dcs01/active/projects/GTeX/syang/look.version8/GO_enrich/MsigDB/c5.all.v7.0.symbols.gmt')


index = grep('^KEGG_HUNTINGTONS_DISEASE$', names(kegg.sets))
genes <- kegg.sets[[index]]
noblood <- subset(w.samps, tissue != 'Whole Blood')

# go through, extract most significant.
rna <- as.data.frame(matrix(nrow = 1, ncol = 3))
colnames(rna) <- c('gene', 'meta.beta', 'meta.pval')
for(i in 1:length(genes))
{
  gene <- genes[i]
  all.but.blood <- subset(noblood, symbol == gene)
  if(nrow(all.but.blood) > 1){
    m1 <- metamean(all.but.blood$samps, all.but.blood$beta, all.but.blood$SE, comb.random = T)
    meta.beta <- m1$TE.random
    tval <- meta.beta/m1$seTE.random
    
    meta.pval <- 2*pt(abs(tval), m1$df.Q, lower=FALSE) # some of these pvals are still greater than one.....
    new <- c(gene, meta.beta, meta.pval)
    rna <- rbind(rna, new)
  }
}

rna$meta.pval <- as.numeric(rna$meta.pval)
rna <- rna[order(rna$meta.pval),]

# sanity check, these betas and pvals are the same as what you get here:
# '/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_objects/meta.vals.rds'

# so just match genes from here instead!
# '/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_objects/meta.vals.rds'
load('/Volumes/JHPCE/dcs01/active/projects/GTeX/syang/look.version8/R_objects/meta.vals.rds')
```


# make function for checking driver genes
```{r}
get.drivers = function(geneset, kegg.sets, meta.vals){
  index = grep(geneset, names(kegg.sets))
  genes <- kegg.sets[[index]]
  driver.genes = meta.vals[which(meta.vals$symbol %in% genes),] 
  return(driver.genes)
}

pd = get.drivers('KEGG_PARKINSONS_DISEASE', kegg.sets, meta.vals)
ad = get.drivers('KEGG_ALZHEIMERS_DISEASE', kegg.sets, meta.vals)
hd = get.drivers('KEGG_HUNTINGTONS_DISEASE', kegg.sets, meta.vals)

pd2 = subset(pd, meta.pval < 0.001)
ad2 = subset(ad, meta.pval < 0.001)
hd2 = subset(hd, meta.pval < 0.001)

head(pd)
head(ad)
head(hd)
```

# make a venn diagram
```{r}
library(VennDiagram)

library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel1")

v = venn.diagram(
  x = list(hd2$gene_id, pd2$gene_id, ad2$gene_id),
  category.names = c("nominally significant HD genes" , "nominally significant  PD genes" , "nominally significant AD genes"),
  filename = NULL,
  output=TRUE,
  
  # Circles
  lwd = 2,
  lty = 'blank',
  fill = myCol,
  
  # # Numbers
  cex = 1.2,
  fontface = "bold",
  
  # # Set names
  cat.cex = 1,
  cat.fontface = "bold",
  ext.pos = list(c(0,0), c(0,0), c(0,0)),
  ext.text = TRUE
  # cat.just = list(c(0,0) , c(1,1) , c(0,2))
)

grid.newpage()
pushViewport(viewport(width=unit(0.8, "npc"), height = unit(0.8, "npc")))
grid.draw(v)
```


# check elk1, RNA-BP, ribosome, mt-part genes
```{r}
elk1 = get.drivers('^SCGGAAGY_ELK1_02$', tft.sets, meta.vals)
rna = get.drivers('^GO_RNA_BINDING$', go.sets, meta.vals)
rib = get.drivers('^KEGG_RIBOSOME$', kegg.sets, meta.vals)
mito = get.drivers('^GO_MITOCHONDRIAL_PART$', go.sets, meta.vals)
```

# check mitorib
```{r}
mtrib = get.drivers('GO_MITOCHONDRIAL_RIBOSOME_ASSEMBLY', go.sets, meta.vals)
```


# pretty table of genes
```{r}
to.show <- rbind(head(elk1, 3), head(rna, 3), head(rib, 3), head(mito, 3))
to.show$meta.beta <- as.numeric(to.show$meta.beta)
to.show$meta.beta <- formatC(to.show$meta.beta, format = "e", digits = 2)
to.show$meta.pval <-ifelse(to.show$meta.pval > 0.001, format(round(to.show$meta.pval, 3), nsmall = 3)
, formatC(to.show$meta.pval, format = "e", digits = 2))
to.show = dplyr::select(to.show, symbol, meta.beta, meta.pval)

library(htmlTable)
htmlTable(to.show,
          header =  c('Gene', '&emsp;&emsp;Meta Effect estimate&emsp;&emsp;', 'Meta Pval'),
          rnames = rep('', nrow(to.show)),
          rgroup = c("ELK1 targets", "RNA-binding genes",
                     "Ribosome genes", 'Mitochondrial part genes'),
          n.rgroup = c(3, 3, 3, 3))


```

# forestplot function:
```{r}
makot.noblood.meta <- function(gene, combined, title = ''){
	# you will save the plot as gene_title
	gene.title <- paste0(gene, title)

	# required packages:
	require(forestplot)
	require(meta)

	# get effects for specific gene:	
	gene.only <- subset(combined, symbol == gene) # 49 tissues --> but might not be same for every tissue!
	gene.only <- gene.only[order(gene.only$t_value, decreasing = T),]

	# get means and upper/lower limits
	means <- gene.only$beta
	upper <- gene.only$beta+gene.only$SE
	lower <- gene.only$beta-gene.only$SE

	# random effects meta-analysis, inverse variance weighted
	# exclude blood from the meta-analysis
	no.blood <- subset(gene.only, tissue != 'Whole Blood')
	
	m1 <- metamean(no.blood$samps, no.blood$beta, no.blood$SE, comb.random = T)
	
	# add meta to vectors
	means <- c(NA, NA, NA, means, NA, m1$TE.random)
	upper <- c(NA, NA, NA, upper, NA, m1$TE.random + m1$seTE.random)
	lower <- c(NA, NA, NA, lower, NA, m1$TE.random - m1$seTE.random)

	# create table text
	text <-cbind(c(paste0("Forestplot for ", gene), NA, "Tissue", gene.only$tissue, NA, 'Summary (Random)'), 
		c(NA, NA, 'N', gene.only$samps, NA, NA), 
		c(NA, NA,"Beta", formatC(gene.only$beta, format = "e", digits = 2), NA, formatC(m1$TE.random, format = "e", digits = 2)))

	# draw forestplot
#	pdf(paste0('/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_code/Cross.tissue.look/forestplots/noblood/', gene.title, '.pdf'), width = 13.3, height = 7.5, onefile = F)
	forestplot(text, means, lower, upper, col=fpColors(box="royalblue",line="darkblue", summary="royalblue"), is.summary=c(TRUE, FALSE, TRUE, rep(FALSE, nrow(gene.only)+1), TRUE))
#	dev.off()
}
```

# ELK1 forestplot
```{r}
load('R_objects/elk1.rds') # 18 tissues where ELK1 was significantly enriched
elk.tiss <- subset(only.sig, tissue %in% elk1$tissue)

make.forestplot.noblood.meta('ELK1', elk.tiss)
```

