# 6.15.2020
# Adapted from Dan Arking
# Do on command line
# Make a table for manuscript

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lattice)
library(splines)
library(survival)
library(knitr)
```
## KEGG Pathways include Parkinsons and Alzheimers
## check for association in UKB
## Generate combined parkinsons and dementia (Neurodisease)
### check cases in each race group


```{r data, echo=FALSE, cache=TRUE}
data<-readRDS("/dcl01/arking/data/static/UKBiobank/GWAS_data/pheno/ukbPheno_05052020.rds")
# data<-readRDS("C:/Users/darking/Desktop/UKB/phenotypes/ukbPheno_05052020.rds")
##include ethnicity
data$race<-fct_collapse(data$ethnicity,
                         White = c("White","British","Irish","Any other white background"),
                         Mixed = c("Mixed", "White and Black African", "White and Asian","Any other mixed background","White and Black Caribbean"),
                         Asian = c("Asian or Asian British","Indian","Pakistani","Bangladeshi","Any other Asian background"),
                        Black = c("Black or Black British","Caribbean","African","Any other Black background"),
                        Chinese =c("Chinese"),
                        Other = c("Other ethnic group"),
                        Missing = c("Prefer not to answer","Do not know")
                        )
#combine dementia and parkinsons
data <- data %>% mutate(Neurodisease=ifelse((Parkinson==1 | Dementia==1),1,0), incNeurodisease=ifelse((incParkinson==1 | incDementia==1),1,0), prvNeurodisease=ifelse((prvParkinson==1 | prvDementia==1),1,0),ttNeurodisease=pmin(ttDementia,ttParkinson))
#remove as incident is either disease is prevalent
data<- data %>% mutate(incNeurodisease=ifelse(prvNeurodisease==1,NA,incNeurodisease),ttNeurodisease=ifelse(prvNeurodisease==1,NA,ttNeurodisease))
  
t0<-data %>% filter(!(is.na(mtDNA_CN1))) %>% group_by(Neurodisease,race) %>% summarise(Count=n())
t0    
clean.data<-data %>% filter(race =="White" & used.in.pca.calculation==1)
```

## Focus on European ancestry, since all other groups <10 cases
## Only use unrelated individuals (used.in.pca.calculation==1)
## mtDNA_CN1 is corrected for ns(age,df=2) + sex
## mtDNA_CN2 is corrected for ns(age,df=2) + sex + Lymph + Platelet + Mono + NucRBC + Neutrophil + Baso + Eos
### cannot use too many covariates due to small number of cases
### Model 1 = age + sex + mtDNA_CN1
### Model 1ex = age + sex + mtDNA_CN1 (exluding cell type count outliers)
### Model 2 = age + sex + mtDNA_CN2 (exluding cell type count outliers)
### Model 3 = model 2 + SmokingStatus
### Model 4 = model 3 + center
### Model 5 = model 4 + year + month
```{r, echo=FALSE}
log_model<-function(pheno,dataset) { 
  pheno.model<-paste0("prv",pheno)
  model1<-c("age","sex","mtDNA_CN1")
  model2<-c("age","sex","mtDNA_CN2")
  model3<-c(model2,"SmokingStatus")
  model4<-c(model3,"Center")
  #model5<-c(model4,"year","month")
  lm1<-glm(reformulate(model1,pheno.model),data=dataset,family="binomial")
  lm1ex<-glm(reformulate(model1,pheno.model),data=subset(dataset,!(is.na(mtDNA_CN2))),family="binomial")
  lm2<-glm(reformulate(model2,pheno.model),data=dataset,family="binomial")
  lm3<-glm(reformulate(model3,pheno.model),data=dataset,family="binomial")
  lm4<-glm(reformulate(model4,pheno.model),data=dataset,family="binomial")
  #lm5<-glm(reformulate(model5,pheno.model),data=dataset,family="binomial")
  
  #results
  m1<-c(format(coef(summary(lm1))["mtDNA_CN1",],digits=3))
  m1ex<-c(format(coef(summary(lm1ex))["mtDNA_CN1",],digits=3))
  m2<-c(format(coef(summary(lm2))["mtDNA_CN2",],digits=3))
  m3<-c(format(coef(summary(lm3))["mtDNA_CN2",],digits=3))
  m4<-c(format(coef(summary(lm4))["mtDNA_CN2",],digits=3))
  #m5<-c(format(coef(summary(lm5))["mtDNA_CN2",],digits=3))
  
  out<-as.data.frame(rbind(m1,m1ex,m2,m3,m4))
  #out<-as.data.frame(rbind(m1,m1ex,m2,m3,m4,m5))
  return(out)
}
cox_model<-function(pheno,dataset) { 
  pheno.model<-paste0("Surv(tt",pheno,",inc",pheno,")")
  model1<-c("age","sex","mtDNA_CN1")
  model2<-c("age","sex","mtDNA_CN2")
  model3<-c(model2,"SmokingStatus")
  model4<-c(model3,"Center")
  #model5<-c(model4,"year","month")
  lm1<-coxph(reformulate(model1,pheno.model),data=dataset)
  print(summary(lm1))
  lm1ex<-coxph(reformulate(model1,pheno.model),data=subset(dataset,!(is.na(mtDNA_CN2))))
  lm2<-coxph(reformulate(model2,pheno.model),data=dataset)
  lm3<-coxph(reformulate(model3,pheno.model),data=dataset)
  lm4<-coxph(reformulate(model4,pheno.model),data=dataset)
  #lm5<-glm(reformulate(model5,pheno.model),data=dataset)
  
  #results
  m1<-c(format(coef(summary(lm1))["mtDNA_CN1",],digits=3))
  m1ex<-c(format(coef(summary(lm1ex))["mtDNA_CN1",],digits=3))
  m2<-c(format(coef(summary(lm2))["mtDNA_CN2",],digits=3))
  m3<-c(format(coef(summary(lm3))["mtDNA_CN2",],digits=3))
  m4<-c(format(coef(summary(lm4))["mtDNA_CN2",],digits=3))
  #m5<-c(format(coef(summary(lm5))["mtDNA_CN2",],digits=3))
  
  out<-as.data.frame(rbind(m1,m1ex,m2,m3,m4))
  #out<-as.data.frame(rbind(m1,m1ex,m2,m3,m4,m5))
  return(out)
}
```
## Parkinsons
```{r Parkinsons, echo=FALSE}
t1<-clean.data %>% filter(!(is.na(mtDNA_CN1))) %>% group_by(Parkinson, prvParkinson) %>% summarise(Count=n())
t1
kable(log_model("Parkinson",clean.data),caption="Prevalent Disease")
kable(cox_model("Parkinson",clean.data),caption="Incident Disease")

pd = cox_model("Parkinson",clean.data)
save(pd, file = '/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_objects/UKB/pd.rds')
```
## Alzheimers
### too few to run prevalent disease
```{r Alzh, echo=FALSE}
t2<-clean.data %>% filter(!(is.na(mtDNA_CN1))) %>% group_by(Alzh, prvAlzh) %>% summarise(Count=n())
t2
kable(cox_model("Alzh",clean.data),caption="Incident Disease")

ad = cox_model("Alzh",clean.data)
save(ad, file = '/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_objects/UKB/ad.rds')
```
## Dementia (includes Alzh)
```{r dementia, echo=FALSE}
t3<-clean.data %>% filter(!(is.na(mtDNA_CN1))) %>% group_by(Dementia, prvDementia) %>% summarise(Count=n())
t3
kable(log_model("Dementia",clean.data),caption="Prevalent Disease")
kable(cox_model("Dementia",clean.data),caption="Incident Disease")
kable(cox_model("Dementia",clean.data %>% filter(Alzh==0)),caption="Incident Disease (No Alzh)")

dem = cox_model("Dementia",clean.data %>% filter(Alzh==0))
save(dem, file = '/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_objects/UKB/dem.rds')
```

### strong effect for Alzh, but probably true for non-Alzh dementia as well
## check overlap between Dementia and Parkinsons
```{r, echo=FALSE}
t4<-clean.data %>% filter(!(is.na(mtDNA_CN1))) %>% group_by(Dementia, Parkinson) %>% summarise(Count=n())
t4
t5<-clean.data %>% filter(!(is.na(mtDNA_CN1))) %>% group_by(Neurodisease,prvNeurodisease) %>% summarise(Count=n())
t5
kable(log_model("Neurodisease",clean.data),caption="Prevalent Disease")
kable(cox_model("Neurodisease",clean.data),caption="Incident Disease")
nd = cox_model("Neurodisease",clean.data)
save(nd, file = '/dcs01/arking/arkinglab/active/projects/GTeX/syang/look.version8/R_objects/UKB/nd.rds')

table(clean.data$incParkinson, clean.data$incAlzh)
```

```{r, echo=FALSE, eval=FALSE}
summary(coxph(Surv(ttNeurodisease,incNeurodisease)~mtDNA_CN1+sex+ns(age,df=2)+SmokingStatus,data=clean.data))
summary(glm(prvNeurodisease~mtDNA_CN1+sex+ns(age,df=2)+SmokingStatus,data=clean.data,family="binomial"))
```

```{r, echo=FALSE, eval=FALSE}
summary(coxph(Surv(ttNeurodisease,incNeurodisease)~mtDNA_CN1+sex+ns(age,df=2)+SmokingStatus,data=clean.data))
summary(glm(prvNeurodisease~mtDNA_CN1+sex+ns(age,df=2)+SmokingStatus,data=clean.data,family="binomial"))

summary(coxph(Surv(ttNeurodisease,incNeurodisease)~mtDNA_CN1+sex+ns(age,df=2),data=clean.data))

clean.data$ttAlz %>% median
clean.data$ttParkinson %>% median
clean.data$ttDementia %>% median
clean.data$ttNeurodisease %>% median
```

